Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making

### 2020-09-15 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
cambios menores en motivation

\subsection{Motivation}
Intelligent agents often face situations in which an action must be chosen in the presence of uncertain conditions; this means that an outcome will be observed according to some probability distribution given the action chosen by the agent. In many real-world applications, the agent doesn't have access to all of the parameters, or even the distribution itself, required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then she could attempt to discover these relations and \textit{use them} in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning \citep{sutton1998reinforcement}.

Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data. For example, current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although they have good performance in the task that were supposed to solve, they cannot explain \textit{why} a specific trajectory was chosen by the algorithm. This is highly relevant in real-world applications, e.g. self-driving cars, where it is very important to understand why an accident happened.

@@

\subsection{Aim}
Our goal is to establish the foundations for decision making algorithms that rely on the existence of causal relations even though such relations are not previously known or fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying, gonzalez2018playing}. We provide here an easily implementable decision making criterion which we build over known results from von Neumman and Morgenstern, Savage and Pearl. With our result we expect to highlight the importance and applicability of causal relations in decision making and the wide variety of possible applications.

>>>>

\subsection{Motivation}
Intelligent agents often face situations in which an action must be chosen in the presence of uncertain conditions; this means that an outcome will be observed according to some probability distribution given the action chosen by the agent. In many real-world applications, the agent doesn't have access to all of the parameters, or even the distribution itself, required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then she could attempt to discover these relations and \textit{use them} in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning \citep{sutton1998reinforcement}. As it is mentioned by \cite{pearl2009causality}, avoiding going to the doctor's office would be the best decision in order to avoid getting sick.

Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data. For example, current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although they have good performance in the task that were supposed to solve, such methods were not designed to explain \textit{why} a specific trajectory was chosen. This is highly relevant in real-world applications, e.g. self-driving cars, where it is very important to understand, for example, why an accident happened.

@@

\subsection{Aim}
Our goal is to establish the foundations for decision making algorithms that rely on causal relations even though such relations are not previously known nor fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying,gonzalez2018playing}. We provide here an easily implementable decision making criterion which we build over known results from von Neumman and Morgenstern, Savage and Pearl. With our result we expect to highlight the importance and applicability of causal relations in decision making and the wide variety of possible applications.
